\chapter*{Conclusion}
\addcontentsline{toc}{chapter}{Conclusion}
\chaptermark{Conclusion}
\markboth{Conclusion}{Conclusion}

We will first discuss the time evolution, energy levels, and spaghetti diagrams.

The time evolution across all systems in
\cref{fig:time-evolution-2,fig:time-evolution-3,fig:time-evolution-4,fig:time-evolution-5}
looks just as one expects. Motion is generally oscillatory like for the
two-level atom, but as $N$ increases higher frequencies are available for the
system evolution and manifest like the wiggles in \cref{fig:time-evolution-5}.

The exponential fits at $\eta = 10$ in
\cref{fig:exponential-fit-2,fig:exponential-fit-3,fig:exponential-fit-4,fig:exponential-fit-5}
are reasonable, but the discrepancies from the true envelopes serve as a
reminder that the spin relaxation plots to follow are to be taken with a grain
of salt.

The energy levels across
\cref{fig:energy-levels-2,fig:energy-levels-3,fig:energy-levels-4,fig:energy-levels-5}
depend on whether $N$ is even or odd, as expected from the solution of the full
transverse-field Ising model, but generally show that the energy levels for
$\hat{\ham}_x$ as $g \to 0$ are far more degenerate than those for
$\hat{\ham}_z$ as $g \to 1$. This offers an explanation of the trends for the
relaxation rates. Take \cref{fig:energy-levels-5} for example. For small $g$,
roughly half of the frequencies are zero and half of the frequencies are two
(neglecting the lowest eigenvalue below $-4$). For large $g$, the energies are
divided up more. For large $\eta$, \cref{eq:gamma-nondim} is like
$\tilde{\omega}^3$, and we see that the small $g$ case has the larger sum of
cubes. This difference in the energy levels for different $g$ may explain the
general shape of the limiting rates discussed below as $\eta \to \infty$. The
effect only arises as $N$ grows because there are an insufficient number of
eigenvalues to split at small $N$. At $N = 2$, the maximal similarity between
the different sides of the energy level diagram may explain the comparable decay
rates found as well.

The spaghetti diagrams
(\cref{fig:spin-spectrum-2,fig:spin-spectrum-3,fig:spin-spectrum-4,fig:spin-spectrum-5})
are difficult to decipher. Again, since the eigenoperators $\opr{V}_i$ of
$\sopr{L}$ are not even states, it is difficult to ascribe meaning to a
particular eigenvalue (noodle). Looking to \cref{fig:spin-spectrum-3}, the
overlaid spin relaxation rate for $\eta = 10$ starts to blend in with the
spaghetti near $g = 1$, but behaves differently elsewhere. Here and in later
spaghetti diagrams, the overlaid rate appears to rise up as $g \to 0$ as it
follows all of the other noodles, yet tends to remain within about a factor of
ten away from the value at $g = 1$. The large magnitude of the spaghetti for $N
= 4$ appears to be consistent with the even $N$ behavior of the transverse-field
Ising model. The full diagram for $N = 6$ could not be computed, but its largest
value at $g = 0.01$ was found to be $586.9$, which is similar to the general
magnitude for $N = 4$.

Overall, the single-spin relaxation plots
(\cref{fig:spin-relaxation-2,fig:spin-relaxation-3,fig:spin-relaxation-4,fig:spin-relaxation-5})
demonstrate that relaxation rates do depend on the system being considered at
all temperatures. It should be reiterated that these plots show the
dimensionless rates, so while the difference between $g = 0$ and $g = 1$ at high
temperature in \cref{fig:spin-relaxation-4} is about a factor of six, this is a
multiplier for an overall rate $\tilde{\gamma}_0$ that may be quite large for
any practical purposes. Given that caveat, we identify two general types of
behavior depending on temperature. Across all numbers of spins, the
high-temperature relaxation rates approach a fixed shape that varies for each
$N$:
\begin{itemize}
  \item For $N = 2$ there is a minimum relaxation rate around $g = 0.65$.
  \item For $N = 3$ there is a maximum relaxation rate around $g = 0.40$.
  \item For $N = 4$ there is a remnant of a peak but the rate generally
    decreases with increasing $g$.
  \item For $N = 5$ there is a maximum relaxation rate around $g = 0.45$.
\end{itemize}

The story at low temperature is a bit different.
\begin{itemize}
  \item For $N = 2$ the rates appear to quickly reach a parabola-like shape for
    $\eta \approx 10$. Afterwards, the fitting becomes unstable as demonstrated
    by the reversal of the temperature dependence for the rates and the large
    error ribbons in the last plot.
  \item For $N = 3$ rates decrease somewhat linearly until $g = 0.5$ and then
    level out for $\eta \approx 10$. The later plots at lower temperature
    demonstrate the convergence of the rates for small $g$ but show instability
    in the fit for $g$ near $0.5$.
  \item For $N = 4$ the convergence to the previous line-flat shape is quicker
    and the transition from low to high $g$ across $g = 0.5$ is smoothed. The
    fit is stable and the shape remains for very low temperatures in the last
    plot.
  \item For $N = 5$ the flat part of the curve has lifted to reveal a parabola
    shape with a minimum at about $g = 0.45$.
\end{itemize}

Again, how much does varying the parameter $g$ change the relaxation rate of
$\ev{\pauli_1^z}$? In short, the answer is by at most a factor of ten.

Given the deviation of the single-spin relaxation rate from the rest of the
spaghetti, it is unclear how well these results generalize to other systems or
observables. The large $N$ limit remains to be explored, but is computationally
intractable with these methods. It would be interesting to consider a wide range
of other observables and initial states to better understand the filtering of
the spaghetti diagrams to produce a single effective decay rate for the
observable. Future work could also look at different systems. Given our
normalization procedure, this could lead to a more systematic exploration that
looks at arbitrary Hamiltonians. Perhaps the theory of random matrices could
provide statistical predictions for normalized decay rates, since statistics for
the eigenvalue spacings $\omega$ are well-understood.

